### 2018-05-30 5_eval.tex
Add Acknowledgments.
Final

Consider two equivalent functions in C (left) and the \porthos[1] input language (right) represented in Figure~\ref{ex:test-both-cf}.
The functions contain three nested \texttt{while}-loops and thus can serve as an illustration of differences in the program compilation and unrolling processes between \porthos[1] and \porthos[2].

@@

The functions are analysed for reachability of the final state by \porthos[2] and modified version \porthos[1] which is able to render the program AST to an event-based control-flow graph (for that, the control-flow instructions \texttt{if-then-else}, \texttt{while} and \texttt{sequence} are expanded recursively and the head and the tails of each instruction are bound by the method processing the parent).
The graph generation is performed via the open-source library \texttt{Graphviz}~\cite{ellson2001graphviz}).
%Each event of the control-flow graph contains the unique number (generated by the \texttt{hashCode} method) in curly brackets below the value, which is necessary for correct displaying the graph.

@@

As another example of program liable for analysis consider the Dekker's algorithm for mutual exclusion of two processes, originally described by Dijkstra~\cite{dijkstra1962over}; the program is presented in Appendix~\ref{apx:dekker}.
For testing \porthos[1], we used the same file in old \porthos{} input language (\texttt{dekker.pts}), which was used in evaluation tests in the original paper~\cite{Porthos17a}.
For performing tests, \porthos[2] was slightly modified in the branch \textit{ay/test-timing}%
\footnote{The modifications were pushed into the \porthos[1] repository \url{github.com/hernanponcedeleon/Dat3M}} %
%
for printing detailed timing information and statistics about number of events (commit f8a6c97f) and writing this information in strict \textit{json} format so that it can be easily processed later by a script-analyser (commit 48ba7161).
The version of \porthos[2] used for the tests is actual up to commit 01dde6b1 of the \porthos[2] repository%
\footnote{\porthos[2] repository: \url{https://github.com/ajuszkowski/PorthosC}}. %
%
Both \porthos[1] and \porthos[2] are run in the state reachability mode for \texttt{TSO} memory model (see sample output of \porthos[2] in Appendix~\ref{apx:output}).

@@

%The performance of the encoding with the new unrolling algorithm used by \porthos[2] is approx. $1.6$ times worse than performance with the old unrolling scheme.
An example with detailed execution time information for \porthos[2] can be found in Appendix~\ref{apx:output} (part "\texttt{timers}" of the output).
Note that the time spent for the interpretation (0.222 sec) and unrolling (0.088 sec) stages remains negligible comparing to the full execution time (40.679 sec).
Therefore, we conclude that the new architecture implies no performance overhead comparing to the previous version of the tool.
%The time spent by the SMT-solver while considering the SMT-formula encoded by \porthos[2] is 0.397 sec, and in case of \porthos[1] the time was 1.291 sec.
%The number of events in the event-flow graph of \porthos[2] is 82 events (among them 59 memory events), and the number of events encoded by \porthos[1] is 95 (among them 51 memory events).
%Even though the number of memory events processed by \porthos[1] was less, the time of solving the result SMT-formula was significantly more.

@@

For evaluating the tool working in the portability analysis mode, we used the same files \texttt{Dekker.c} and \texttt{Dekker.pts} tested for the state-portability from TSO to SC: %TODO: Say sth about state-portability
%!!! \textit{\textbf{TODO: change the main class name to PorthosC }} !!!
As the portability analysis requires compiling the program under two memory models, the overall program domain encoding time 65.693 sec is almost double as the same time in reachability analysis mode (\porthos[1] shows program domain encoding time 41.027 sec with the bound $k_1=3$).

@@

\end{lstlisting}
\caption{Example of interpretation of a code with an arbitrary control-flow}
\label{ex:cfg-arbitrary}
\end{figure}

@@

%The revised design of \porthos{} tool includes the following new features:
Apart from redesigning the tool architecture and extending the input language, we added support for basic constructions used in \textit{Kernel litmus tests}~\cite{alglave2018frightening}.
Although, current version of \porthos[2] only supports some basic macros of the Linux kernel (such as \texttt{READ\_ONCE} and \texttt{WRITE\_ONCE} for atomic load and store with relaxed memory ordering) as the support for kernel-specific memory barriers goes out of the scope of current thesis.
Note that, comparing to the \porthos[1], adding support for new functions in \porthos[2] does not require changing the input language grammar, this is carried by invocation hooking mechanism described in Section~\ref{ch:impl:proc:x-compiler}.

As an example, consider the SB litmus test for the Linux kernel (which is similar to the one presented in Figure~\ref{simple_wmm_x86}):
\begin{lstlisting}
{ int x = 0; int y = 0;}

@@

exists(x == 0 && y == 0)
\end{lstlisting}

When verifying this litmus test by \porthos[2] in the state reachability mode, the test passes for TSO memory model and fails for SC model.

>>>>

Consider two equivalent functions \texttt{t0} in Figure~\ref{ex:test-both-cf}: in the C language (left) and in the \porthos[1] input language (right).
The functions contain three nested \texttt{while}-loops and thus can serve as an illustration of differences in the program compilation and unrolling processes between \porthos[1] and \porthos[2].

@@

The functions are analysed for reachability of the final state by \porthos[2] and modified version \porthos[1] which is able to render the program AST to an event-based control-flow graph (for that, the branching statement \texttt{if-then-else}, the \texttt{while} loop and the \texttt{sequence} statement are expanded recursively and the head and the tails of each instruction are bound by the method processing the parent).
The graph generation is performed via the open-source library \texttt{Graphviz}~\cite{ellson2001graphviz}).
%Each event of the control-flow graph contains the unique number (generated by the \texttt{hashCode} method) in curly brackets below the value, which is necessary for correct displaying the graph.

@@

As another example of program liable for analysis consider the Dekker's algorithm for mutual exclusion of two processes, originally described by Dijkstra~\cite{dijkstra1962over}; the program is presented in Appendix~\ref{apx:dekker}.
For testing \porthos[1], we used the same file in old \porthos{} input language (\texttt{dekker.pts}), which was used in evaluation tests in the original paper~\cite{Porthos17a}.
For performing tests, \porthos[2] was slightly modified in the branch \textit{ay/test-timing}%
\footnote{The modifications were pushed into the \porthos[1] repository \url{github.com/hernanponcedeleon/Dat3M}} %
%
for printing detailed timing information and statistics about number of events (commit f8a6c97f) and writing this information in strict \textit{json} format so that it can be easily processed later by a script that analyses and prints the results (commit 48ba7161).
The version of \porthos[2] used for the tests is actual up to commit 01dde6b1 of the \porthos[2] repository%
\footnote{\porthos[2] repository: \url{https://github.com/ajuszkowski/PorthosC}}. %
%
Both \porthos[1] and \porthos[2] are run in the state reachability mode for TSO memory model (see sample output of \porthos[2] in Appendix~\ref{apx:output}).

@@

%The performance of the encoding with the new unrolling algorithm used by \porthos[2] is approx. $1.6$ times worse than performance with the old unrolling scheme.
An example with detailed execution time information for \porthos[2] can be found in Appendix~\ref{apx:output} (part `\texttt{timers}' of the output).
Note that the time spent for the interpretation (0.222 sec) and unrolling (0.088 sec) stages remains negligible comparing to the full execution time (40.679 sec).
Therefore, we conclude that the new architecture implies no performance overhead comparing to the previous version of the tool.
%The time spent by the SMT-solver while considering the SMT-formula encoded by \porthos[2] is 0.397 sec, and in case of \porthos[1] the time was 1.291 sec.
%The number of events in the event-flow graph of \porthos[2] is 82 events (among them 59 memory events), and the number of events encoded by \porthos[1] is 95 (among them 51 memory events).
%Even though the number of memory events processed by \porthos[1] was less, the time of solving the result SMT-formula was significantly more.

@@

For evaluating the tool working in the portability analysis mode, we used the same files \texttt{Dekker.c} and \texttt{Dekker.pts} tested for the state-portability from TSO to SC: %TODO: Say sth about state-portability
%!!! \textit{\textbf{TODO: change the main class name to PorthosC }} !!!
As the portability analysis requires compiling the program under two memory models, the overall program domain encoding time 72.744 sec with unrolling bound $k_2=27$ is almost double as the same time in reachability analysis mode (\porthos[1] shows program domain encoding time 41.027 sec with the bound $k_1=3$).

@@

\end{lstlisting}
\caption{Example: A litmus test in C with an arbitrary control-flow}
\label{ex:cfg-arbitrary}
\end{figure}

@@

\begin{figure}[t]
\begin{lstlisting}
{ int x = 0; int y = 0;}

@@

exists(x == 0 && y == 0)
\end{lstlisting}
\caption{SB litmus test for the Linux kernel memory model}
\label{ex:sb:kernel}
\end{figure}

%The revised design of \porthos{} tool includes the following new features:
Apart from redesigning the tool architecture and extending the input language, we added support for basic constructions used in \textit{Kernel litmus tests}~\cite{alglave2018frightening}.
Although, current version of \porthos[2] only supports some basic macros of the Linux kernel (such as \lstinline{READ_ONCE} and \lstinline{WRITE_ONCE} for atomic load and store with relaxed memory ordering) as the support for kernel-specific memory barriers goes out of the scope of current thesis.
Note that, comparing to the \porthos[1], adding support for new functions in \porthos[2] does not require changing the input language grammar, this is carried by invocation hooking mechanism described in Section~\ref{ch:impl:proc:x-compiler}.

As an example, consider the Store Buffering litmus test for the Linux kernel in Figure~\ref{ex:sb:kernel} (which is similar to the one presented in Figure~\ref{simple_wmm_x86}).
When verifying this litmus test by \porthos[2] in the state reachability mode, the test passes for TSO memory model and fails for SC model.

### 2018-05-30 90_summary.tex
Final

\item The wide range of existing litmus tests may require \porthos[2] to \textit{support new input languages}.
The first candidates for being supported are the LISA language (Litmus Instruction Set Architecture)~\cite{alglave2016syntax} and assembly languages of most common architectures (x86, POWER, Alpha, etc.).
Note that the programs in low-level languages can be easily compiled to the \xgraph{} representation by existing compiler architecture since low-level assembly-like languages can be considered as the subset of the C language, that restricts the set of non-linear control-flow instructions to the conditional and unconditional jumps (which are supported by the current \xgraph{} compiler of \porthos[2]).
%Note that adding support for low-level languages requires adding the new \xgraph{} compiler that will compile the low-level syntax tree (with non-conditional jumps as the only )

>>>>

\item The wide range of existing litmus tests may require \porthos[2] to \textit{support new input languages}.
The first candidates for being supported are the LISA language (Litmus Instruction Set Architecture)~\cite{alglave2016syntax} and assembly languages of most common architectures (x86, Power, Alpha, etc.).
Note that the programs in low-level languages can be easily compiled to the \xgraph{} representation by existing compiler architecture since low-level assembly-like languages can be considered as the subset of the C language, that restricts the set of non-linear control-flow instructions to the conditional and unconditional jumps (which are supported by the current \xgraph{} compiler of \porthos[2]).
%Note that adding support for low-level languages requires adding the new \xgraph{} compiler that will compile the low-level syntax tree (with non-conditional jumps as the only )
